mruby Bytecode Loader Using Bluetooth in Multi-VM Environment

### 2016-03-19 EMSOFT2016/thesis.tex
update

%
% File: thesis.tex
% Created: 月 11 30 08:00 午後 2015 東京 (標準時)
% Last Change: 月 11 30 08:00 午後 2015 東京 (標準時)
%
\input{dummy/dummy.tex}
\documentclass[conference,compsoc]{IEEEtran}
\usepackage[dvipdfmx]{graphicx}
\usepackage{amssymb}
\usepackage{enumerate,cite,url}
\usepackage{listings,jlisting}
\lstset{%
    language={c},%
    basicstyle={\small},%
    identifierstyle={\small},%
    commentstyle={\small\itshape},%
    keywordstyle={\small},%\bfseries},%
    ndkeywordstyle={\small},%
    stringstyle={\small\it},
    frame={tb},
    breaklines=true,
    columns=[l]{fullflexible},%
    numbers=left,%
    xrightmargin=0zw,%
    xleftmargin=3zw,%
    numberstyle={\scriptsize},%
    stepnumber=1,
    numbersep=1zw,%
    %lineskip=-0.5ex%
}

\title{mruby Bytecode Loader Using Bluetooth in Multi-VM Environment}
\author{Takuro Yamamoto}
\date{\today}

\begin{document}
\maketitle
\begin{abstract}
In recent years, the productivity of embedded systems has become a problem due to their complexity and large-scale.
For the purpose of improving the productivity for embedded software development, the mruby on TECS framework has been proposed that is applied mruby (Lightweight Ruby) and supports component-based development.
In the current mruby on TECS, the mruby programs have to be compiled and linked every time the programs are modified because the mruby bytecodes are incorporated in the platform.
Moreover, while the framework supports multi-VM, developers need to be familiar with the functions of real-time operating systems to effectively execute multiple mruby programs in concurrent or/and parallel.
To improve the development efficiency, this paper proposes an mruby bytecode loader using Bluetooth as an extension of mruby on TECS.
The loader executes two mruby bytecodes, mruby application bytecode and mruby library bytecode.
mruby application bytecode modified frequently is sent from a host to a target device by developers.
mruby library bytecode modified infrequently is preserved beforehand with the platform in a storage/ROM device at the time of the first compilation.
In addition, multiple mruby programs cooperatively run in the proposed framework.
A RiteVM sceduler makes multitasking processing more easy-to-use than that of mruby on TECS.
Synchronization of starting multiple tasks is also implemented with Eventflag. 
Experimental results demonstrate the advantages of the proposed framework.
\end{abstract}
\section{Introduction}
In these days, embedded systems have been required the high-quality and the high-performance.
Due to this trend, the complexity of embedded systems also increases and the scale is larger.
For example, IoT (Internet of Things) applications.
The low production cost and the short developing period of time are also needed.

@@

%This paper proposes two additional features of mruby on TECS, mruby bytecode loader using Bluetooth and RiteVM sceduler.
mruby on TECS has several problems at present.
One of the problems is low software development efficiency.
mruby on TECS only supports a storage/ROM device in the platform for LEGO MINDSTORMS EV3 \cite{par:EV3} to load mruby programs.
For example, an SD card should be inserted and pulled out repeatedly, or a ROM should be rewritten if mruby programs are modified.
Developers should also restart an OS on a target device.
%Moreover, in the current system if developers attempt to run multiple tasks with the same priority, a task with first execution only runs and all other tasks do not run unless developers call the OS's function.
%In addition developers must prepare as many VMs as the number of tasks because one RiteVM supports only one task.
Moreover, although mruby on TECS has supported the multi-VM, developers need to call OS's function in order to execute multiple tasks.
These problem are a big burden on developers.

This paper proposes an extended framework of mruby on TECS: an mruby bytecode loader using Bluetooth and a RiteVM scheduler for fairly executing mruby programs.
To improve the development efficiency, the mruby bytecode loader using Bluetooth enables developers to write the platform in a strage device only once at the beginning and to transfer mruby application codes from a host to a target device.
The RiteVM scheduler enables developers to run multitasking programs more easily than mruby on TECS.
%Specifically, a periodic handler of switching tasks is implemented.
%This paper evaluates the proposed framework in term of data transfer rate, overhead of periodic handler and execution time.

{\bf Contribution}: The proposed framework gives the contribution in the following points.
\begin{enumerate}
\item {\bf To improve the software development efficiency:} Developers do not need to rewrite a storage/ROM device and also to restart an OS.
The loader supports the countinuous loading, which saves the Bluetooth set-up time.
Therefore, the mruby bytecode loader using Bluetooth helps developers develop software.
\item {\bf To effectively execute multiple mruby programs in concurrent or/and parallel:} Developers can execute multiple tasks without the knowledge of an RTOS because the RiteVM scheduler switches tasks cyclically.
In addition, synchronization of multiple tasks is also implemented.
\item {\bf To focus on the benefits of component-based developments:} This paper shows the specific examples for the benefits of component-based development. 
\end{enumerate}

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/proposed.pdf}
    \caption{System Model}
    \label{fig:proposed}
\end{figure}

@@

\subsection{mruby}
\label{sec:mruby}
mruby is the light-weight implementation of Ruby programming language complying to part of the ISO standard.

Ruby is a object-oriented script language \cite{url:Ruby}.
As the main feature, Ruby is easy-to-use and easy-to-read due to its simple grammar.
Ruby can run a program with fewer code lines than C language.
Ruby improves the productivity of a software development owing to not only simple grammar also object-oriented functions such as classes and methods, exceptions, and garbage collection.

mruby is suitable for embedded systems because of faster execution with less amount of resources and takes over the usability and readability of Ruby.
In addition, VM (Virtual Machine) mechanism is used in mruby, therefore mruby programs can run on any operating system as long as VM is implemented.
A RiteVM is a VM in mruby, that runs mruby programs.
The RiteVM mechanism is shown in Figure \ref{fig:mruby}.
The mruby compiler translates an mruby code into a bytecode, which is an intermediate code that can be interpreted by a RiteVM.
The bytecodes can run on a Rite VM, and thus mruby programs can be executed on any target devices if only RiteVMs are implemented.
\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/mruby.pdf}
    \caption{Mechanism of mruby/RiteVM}
    \label{fig:mruby}
\end{figure}

\subsection{TECS}
\label{sec:TECS}
TECS (TOPPERS Embedded Component System) is a component system suitable for embedded systems.
TECS helps decrease difficulty because the generated component diagram can visualize the structure of whole software.
It can also help increase productivity and reduce development duplication because a common part of software is regarded as a component.

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/component_diagram.pdf}
    \caption{Component Diagram}
    \label{fig:component}
\end{figure}

\subsubsection{Component Description}\mbox{}\\

The description of a component in TECS is divided into three parts, {\it signature}, {\it celltype}, and build description.
TECS code is written in .cdl (component description language) file.
The component description is mentioned with an example shown in Figure \ref{fig:component} as follow.

\begin{description}
    \item[{\bf Signature Description}]\mbox{}\\
        The {\it signature} description defines an interface of a {\it cell}.
        A {\it signature} name is described following the keyword {\it signature}.
        It also has the prefix ``s".
        In this way, a {\it signature} is defined such as sMotor shown in Figure \ref{signature}.
        To make the definition of an interface clear, specifiers such as in and out are used in TECS.
        [in] and [out] represent input and output, respectively.\\
\begin{figure}[t]
\centering
\begin{lstlisting}
signature sMotor{
    int32_t getCounts( void );
    ER resetCounts( void );
    ER setPower( [in]int power );
    ER stop( [in] bool_t brake );
    ER rotate( [in] int degrees, [in] uint32_t speed_abs,
              [in] bool_t blocking );
    void initializePort( [in]int32_t type );
};
\end{lstlisting}
\caption{Signature Description}
\label{signature}
\end{figure}

    \item[{\bf Celltype Description}]\mbox{}\\
        The {\it celltype} description defines {\it entry} ports, {\it call} ports, attributes, and valuables of a {\it celltype}.
        An example of a {\it celltype} description is shown in Figure \ref{celltype}.
        A {\it celltype} name following the keyword {\it celltype} with the prefix ``t" and elements of a {\it celltype} is described.
        To define {\it entry} ports, a {\it signature} such as sMotor, and an {\it entry} port name such as eMotor follow the keyword {\it entry}.
        In the same way, {\it call} ports can be declared.
        Attributes and valuables follow the keyword {\it attr} and {\it var} respectively.\\
\begin{figure}[t]
\centering
\begin{lstlisting}
celltype tCaller{
    call sMotor cMotor;
};
celltype tMotor{
    entry sMotor eMotor;
    attr{
        int32_t attr = 100;
    };
    var{
        int32_t var;
    };
};
\end{lstlisting}
\caption{Celltype Description}
\label{celltype}
\end{figure}

    \item[{\bf Build Description}]\mbox{}\\
        The build description is used to instantiate {\it cell}s and connect {\it cell}s.
        Figure \ref{build} shows an example of a build description.
        A {\it celltype} name such as tMotor, and a {\it cell} name such as Motor follow the keyword {\it cell}.
        To compose {\it cell}s, a {\it call} port, a {\it signature}, a {\it entry} port in order are described.
        In this example, a {\it entry} port eMotor in a {\it cell} Motor is connected to a {\it call} port cMotor in a {\it cell} Caller.\\
\begin{figure}[t]
\centering
\begin{lstlisting}
cell tMotor Motor{
};
cell tCaller Caller{
    cMotor = Motor.eMotor;
};
\end{lstlisting}
\caption{Build Description}
\label{build}
\end{figure}

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/mrubyontecs.pdf}
    \caption{System Model of existing mruby on TECS}
    \label{fig:mrubyontecs}
\end{figure}

\subsubsection{mruby-TECS Bridge}\mbox{}\\

There is a great difference between the execution time of mruby and C language.
According to  \cite{par:mrubyonTECS}, mruby programs are several hundreds times slower than C programs.
The execution of mruby bytecode on a RiteVM is not as efficient as that of C language.
Thus it is difficult to use mruby for all of code.

The use of Ruby on embedded devices provides the benefit of productivity and maintainability due to the ease to use and read.
On the other hands, it is necessary to implement parts of applications in C language in order to manipulate actuators and sensors, and also make a critical section of code run quickly.

Figure \ref{fig:mruby_TECS_bridge} shows an example of use of an mruby-TECS bridge for controlling a motor.
The left side of BridgeMotor belongs to the mruby program.
The right side of BridgeMotor belongs to TECS component.
\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/mruby_TECS_bridge.pdf}
    \caption{mruby-TECS Bridge}
    \label{fig:mruby_TECS_bridge}
\end{figure}

The mruby-TECS bridge generates two things.
One is a {\it celltype} to receive invocation from the mruby program.
The other thing is an mruby class that corresponds to the TECS component specified by the developers to invoke a C function from the mruby program.

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/system_model.pdf}
    \caption{Detailed System Model of the proposed framework}
    \label{fig:system_model}
\end{figure}

\subsection{mruby Bytecode Loader Using Bluetooth}
\label{sec:mruby bytecode loader using Bluetooth}
This section describes an additional functionality of mruby on TECS, mruby bytecode loader using Bluetooth.
In the current system, the platform including mruby bytecodes are saved in a storage/ROM device.
Developers must rewrite the storage/ROM device every time the application programs are modified.
The OS on the target device should be also restarted.
It causes a low development efficiency to repeat that.
An mruby bytecode loader using Bluetooth makes the developers' burden decrease.
%The work that an SD card is in and out needs to be done just once in the beginning.
Developers should just once carry out the work to connect the storage/ROM device and start the OS.

mruby programs are consisted of mruby application and libraries.
mruby application is the main code, which developers should program.
mruby libraries are the codes defining the functions for application such as Ruby classes. 
The mruby bytecodes including mruby application and libraries can be sent and run.
However, it is also wasteful in terms of the size and time to send because libraries are not frequently modified. 
The proposed framework provides the design that only mruby applications are sent and mruby libraries are preserved with the platform in the storage/ROM device beforehand.
Due to this design, RiteVMs can share the mruby libraries.
In additon, a RiteVM can use the own library that other RiteVMs should not use.

@@

% \begin{figure}[t]
%     \centering
%     \includegraphics[width=8cm,clip]{figure/bluetooth_loader.pdf}
%     \caption{Development Flow in mruby bytecode loader using Bluetooth}
%     \label{fig:bluetooth_loader}
% \end{figure}

\subsubsection{Component of RiteVM with mruby bytecode loader using Bluetooth} \mbox{}\\

The proposed framework provides a RiteVM with mruby bytecode loader using Bluetooth as TECS component.
The component is an extension of the RiteVM component, which is described in \cite{par:mrubyonTECS}.
The component plays a role in receiving bytecodes via Bluetooth, and also manages a RiteVM configuration such as automatically generates the bytecode in the build description.
This generated bytecode is prepared beforehand in the storage/ROM device such as mruby libraries, and different from a bytecode transferred with Bluetooth.

Figure \ref{fig:component_bluetooth} shows a component diagram of MrubyTask1 and MrubyBluetooth1 {\it cell}s.
The MrubyTask1 {\it cell} is a componentized task of the RTOS (TOPPERS/HRP2).
TOPPERS/HRP2 is described in \cite{url:HRP2}, \cite{par:hr-tecs}.
The MrubyBluetooth1 is a component of RiteVM with mruby bytecode loader using Bluetooth.
A bytecode in the host is transferred and received at the top of the component.
In this framework, ZMODEM \cite{par:zmodem} is used as a binary transfer protocol.

\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/component_bluetooth.pdf}
    \caption{Component Diagram of mruby bytecode loader using Bluetooth}
    \label{fig:component_bluetooth}
\end{figure}

Figure \ref{fig:control_flow} shows the process of executing mruby program in a component of RiteVM with mruby bytecode loader using Bluetooth such as MrubyBluetooth1.
The concrete main code of tMrubyBluetooth is shown in Figure \ref{maincode_mrubybluetooth}.

First, a pointer of {\it mrb\_state} and {\it mrbc\_context} are initialized.
{\it mrb\_state} is a set of states and global variables used in mruby.

Second, the RiteVM reads the bytecode of mruby libraries.
mruby libraries are a set of Ruby classes such as motor class and sensor class.
For example, motor class defines methods to rotate and stop a motor.
The tMrubyBluetooth {\it cell} has attributes as shown in Figure \ref{celltype_mrubybluetooth}.
The {\it mrubyFile} indicates the program files of mruby libraries.
{\it [omit]} is only used for the TECS generator, thus the attribute, {\it mrubyFile}, does not consume memory.
The {\it irep} is the pointer of the array stored the bytecode of mruby libraries.
In short, the bytecode of mruby libraries is stored as an attribute of the component when compiling for the first time.

Third, the RiteVM reads the bytecode of the mruby application transferred with Bluetooth.
The mruby application bytecode is stored in an array of type uint8\_t such as {\it irepApp}, which is a tMrubyBluetooth variable as shown in \ref{celltype_mrubybluetooth}.
The array is different from that of holding the mruby library bytecode.
Two bytecodes are read separately in the RiteVM.

Finally, the mruby task runs.
When the mruby application is modified, only the bytecode of the modified application should be transferred.
mruby libraries need not be touched because libraries are not normally changed.
\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/control_flow.pdf}
    \caption{Control Flow of mruby bytecode loader using Bluetooth}
    \label{fig:control_flow}
\end{figure}
\begin{figure}[t]
\centering
\begin{lstlisting}
celltype tMrubyBluetooth{
    entry sTaskBody eMrubyBody;
    attr{
      [omit]char_t *mrubyFile;
      char_t *irep = C_EXP("&$cell_global$_irep");
      uint32_t irepAppSize = C_EXP( BUFFER_SIZE );
    };
    var{
        [size_is(irepAppSize)] uint8_t *irepApp;
    };
};
\end{lstlisting}
\caption{Celltype Description for RiteVM with mruby bytecode loader using Bluetooth}
\label{celltype_mrubybluetooth}
\end{figure}
\begin{figure}[t]
\centering
\begin{lstlisting}
void
eMrubyBody_main( CELLIDX idx )
{
  /* Declaration variables */
  mrb_state *mrb;
  mrbc_context *c;
  /* New interpreter instance */
  mrb = mrb_open();
  // Omit: error check for mrb_state
  /* New mruby context */
  c = mrbc_context_new( mrb );
  // Omit: initialization of mruby-TECS bridge
  /* Receive the bytecode via Bluetooth */
  bluetooth_loader( VAR_irepApp );
  /* Load mruby library bytecode and run */
  mrb_load_irep_cxt( mrb, ATTR_irep, c );
  /* Load mruby application bytecode and run */
  mrb_load_irep_cxt( mrb, VAR_irepApp, c );
  if ( mrb->exc ) {
    /* Failure to execute */
    mrb_p( mrb, mrb_obj_value( mrb->exc ) );
    exit( 0 );
  }
  /* Free mruby context */
  mrbc_context_free( mrb, c );
  /* Free interpreter instance */
  mrb_close( mrb );
}

>>>>

%
% File: thesis.tex
% Created: 月 11 30 08:00 午後 2015 東京 (標準時)
% Last Change: 月 11 30 08:00 午後 2015 東京 (標準時)
%
\input{dummy/dummy.tex}
\documentclass{sig-alternate-05-2015}
%\usepackage[dvipdfmx]{graphicx}
\usepackage{amssymb}
\usepackage{enumerate,cite,url}
\usepackage{listings,jlisting}
\lstset{%
    language={c},%
    basicstyle={\small},%
    identifierstyle={\small},%
    commentstyle={\small\itshape},%
    keywordstyle={\small},%\bfseries},%
    ndkeywordstyle={\small},%
    stringstyle={\small\it},
    frame={tb},
    breaklines=true,
    columns=[l]{fullflexible},%
    numbers=left,%
    xrightmargin=0zw,%
    xleftmargin=3zw,%
    numberstyle={\scriptsize},%
    stepnumber=1,
    numbersep=1zw,%
    %lineskip=-0.5ex%
}

% Copyright
\setcopyright{acmcopyright}

\title{mruby Bytecode Loader Using Bluetooth in Multi-VM Environment}
\numberofauthors{3}
\author{
\alignauthor
Takuro Yamamoto\\
    \affaddr{Graduate School of Engineering Science, Osaka University}
       % \affaddr{1932 Wallamaloo Lane}\\
       % \affaddr{Wallamaloo, New Zealand}\\
       % \email{trovato@corporation.com}
\alignauthor
Hiroshi Oyama\\
    \affaddr{OKUMA Corporation}
       % \affaddr{P.O. Box 1212}\\
       % \affaddr{Dublin, Ohio 43017-6221}\\
       % \email{webmaster@marysville-ohio.com}
\alignauthor
Takuya Azumi\\
    \affaddr{Graduate School of Engineering Science, Osaka University}
       % \affaddr{1 Th{\o}rv{\"a}ld Circle}\\
       % \affaddr{Hekla, Iceland}\\
       % \email{larst@affiliation.org}
}
%\author{Takuro Yamamoto}
%\date{\today}

\begin{document}
\maketitle
\begin{abstract}
In recent years, the productivity of embedded systems has become a problem due to their complexity and large-scale.
For the purpose of improving the productivity for embedded software development, the mruby on TECS framework has been proposed that is applied mruby (Lightweight Ruby) and supports component-based development.
In the current mruby on TECS, the mruby programs have to be compiled and linked every time the programs are modified because the mruby bytecodes are incorporated in the platform.
Moreover, while the framework supports multi-VM, developers need to be familiar with the functions of real-time operating systems to effectively execute multiple mruby programs in concurrent or/and parallel.
To improve the development efficiency, this paper proposes an mruby bytecode loader using Bluetooth as an extension of mruby on TECS.
The loader executes two mruby bytecodes, mruby application bytecode and mruby library bytecode.
mruby application bytecode modified frequently is sent from a host to a target device by developers.
mruby library bytecode modified infrequently is preserved beforehand with the platform in a storage/ROM device at the time of the first compilation.
In addition, multiple mruby programs cooperatively run in the proposed framework.
A RiteVM scheduler makes multitasking processing more easy-to-use than that of mruby on TECS.
Synchronization of starting multiple tasks is also implemented with Eventflag. 
Experimental results demonstrate the advantages of the proposed framework.
\end{abstract}
\section{Introduction}
In these days, embedded systems have been required the high-quality and the high-performance.
Due to this trend, the complexity of embedded systems also increases and the scale is larger.
For example, IoT (Internet of Things) applications.
The low production cost and the short developing period of time are also needed.

@@

%This paper proposes two additional features of mruby on TECS, mruby bytecode loader using Bluetooth and RiteVM scheduler.
mruby on TECS has several problems at present.
One of the problems is low software development efficiency.
mruby on TECS only supports a storage/ROM device to load mruby programs. 
%in the platform for LEGO MINDSTORMS EV3 \cite{par:EV3}.
For example, an SD card should be inserted and pulled out repeatedly, or a ROM should be rewritten if mruby programs are modified.
Developers should also restart an OS on a target device.
%Moreover, in the current system if developers attempt to run multiple tasks with the same priority, a task with first execution only runs and all other tasks do not run unless developers call the OS's function.
%In addition developers must prepare as many VMs as the number of tasks because one RiteVM supports only one task.
Moreover, although mruby on TECS has supported the multi-VM, developers need to call OS's function in order to execute multiple tasks.
These problem are a big burden on developers.

This paper proposes an extended framework of mruby on TECS: an mruby bytecode loader using Bluetooth and a RiteVM scheduler for fairly executing mruby programs.
To improve the development efficiency, the mruby bytecode loader using Bluetooth enables developers to write the platform in a storage device only once at the beginning and to transfer mruby application codes from a host to a target device.
The RiteVM scheduler enables developers to run multitasking programs more easily than mruby on TECS.
%Specifically, a periodic handler of switching tasks is implemented.
%This paper evaluates the proposed framework in term of data transfer rate, overhead of periodic handler and execution time.

{\bf Contribution}: The proposed framework gives the contribution in the following points.
\begin{enumerate}
\item {\bf To improve the software development efficiency:} Developers do not need to rewrite a storage/ROM device and also to restart an OS.
The loader supports the continuous loading, which saves the Bluetooth set-up time.
Therefore, the mruby bytecode loader using Bluetooth helps developers develop software.
\item {\bf To effectively execute multiple mruby programs in concurrent or/and parallel:} Developers can execute multiple tasks without the knowledge of an RTOS because the RiteVM scheduler switches tasks cyclically.
\item {\bf To synchronize the execution of multiple RiteVM tasks:} The proposed framework provides the synchronization of multiple RiteVM tasks (mruby applications).
\item {\bf To focus on the benefits of component-based developments:} This paper shows the specific examples for the benefits of component-based development. 
\end{enumerate}

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/proposed.eps}
    \caption{System Model of the proposed framework}
    \label{fig:proposed}
\end{figure}

@@

\subsection{mruby}
\label{sec:mruby}
mruby is the light-weight implementation of Ruby programming language complying to part of the ISO standard.
Ruby is a object-oriented script language \cite{url:Ruby}.
As the main feature, Ruby is easy-to-use and easy-to-read due to its simple grammar.
Ruby can run a program with fewer code lines than C language.
Ruby improves the productivity of a software development owing to not only simple grammar but also object-oriented functions such as classes and methods, exceptions, and garbage collection.

mruby is suitable for embedded systems because of faster execution with less amount of resources and takes over the usability and readability of Ruby.
In addition, VM (Virtual Machine) mechanism is used in mruby, therefore mruby programs can run on any operating system as long as VM is implemented.
A RiteVM is a VM in mruby, that runs mruby programs.
The RiteVM mechanism is shown in Figure \ref{fig:mruby}.
The mruby compiler translates an mruby code into a bytecode, which is an intermediate code that can be interpreted by a RiteVM.
The bytecodes can run on a Rite VM, and thus mruby programs can be executed on any target devices if only RiteVMs are implemented.
\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/mruby.eps}
    \caption{Mechanism of mruby/RiteVM}
    \label{fig:mruby}
\end{figure}

\subsection{TECS}
\label{sec:TECS}
TECS (TOPPERS Embedded Component System) is a component system suitable for embedded systems.
TECS helps increase productivity and reduce development duplication because the reusability of software is improved.
TECS also helps decrease complexity of the system because the generated component diagram can visualize the structure of whole software.

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/component_diagram.eps}
    \caption{Component Diagram}
    \label{fig:component}
\end{figure}

\subsubsection{Component Description}
The description of a component in TECS is divided into three parts, {\it signature}, {\it celltype}, and build description.
TECS code is written in .cdl (component description language) file.
The component description is mentioned with an example shown in Figure \ref{fig:component} as follow.

\begin{description}
    \item[{\bf Signature Description}]\mbox{}\\
        The {\it signature} description defines an interface of a {\it cell}.
        A {\it signature} name is described following the keyword {\it signature}.
        It also has the prefix ``s".
        In this way, a {\it signature} is defined such as sMotor shown in Figure \ref{signature}.
        To make the definition of an interface clear, specifiers such as in and out are used in TECS.
        [in] and [out] represent input and output, respectively.
\begin{figure}[t]
\centering
\begin{lstlisting}
signature sMotor{
    int32_t getCounts( void );
    ER resetCounts( void );
    ER setPower( [in]int power );
    ER stop( [in] bool_t brake );
    ER rotate( [in] int degrees, [in] uint32_t speed_abs,
              [in] bool_t blocking );
    void initializePort( [in]int32_t type );
};
\end{lstlisting}
\caption{Signature Description}
\label{signature}
\end{figure}
    \item[{\bf Celltype Description}]\mbox{}\\
        The {\it celltype} description defines {\it entry} ports, {\it call} ports, attributes, and valuables of a {\it celltype}.
        An example of a {\it celltype} description is shown in Figure \ref{celltype}.
        A {\it celltype} name following the keyword {\it celltype} with the prefix ``t" and elements of a {\it celltype} is described.
        To define {\it entry} ports, a {\it signature} such as sMotor, and an {\it entry} port name such as eMotor follow the keyword {\it entry}.
        In the same way, {\it call} ports can be defined.
        Attributes and valuables follow the keyword {\it attr} and {\it var} respectively.
\begin{figure}[t]
\centering
\begin{lstlisting}
celltype tCaller{
    call sMotor cMotor;
};
celltype tMotor{
    entry sMotor eMotor;
    attr{
        int32_t attribute;
    };
    var{
        int32_t variable;
    };
};
\end{lstlisting}
\caption{Celltype Description}
\label{celltype}
\end{figure}
    \item[{\bf Build Description}]\mbox{}\\
        The build description is used to instantiate {\it cell}s and connect {\it cell}s.
        Figure \ref{build} shows an example of a build description.
        A {\it celltype} name such as tMotor, and a {\it cell} name such as Motor follow the keyword {\it cell}.
        To compose {\it cell}s, a {\it call} port, a {\it signature}, a {\it entry} port in order are described.
        In this example, a {\it entry} port eMotor in a {\it cell} Motor is connected to a {\it call} port cMotor in a {\it cell} Caller.
\begin{figure}[t]
\centering
\begin{lstlisting}
cell tMotor Motor{
};
cell tCaller Caller{
    cMotor = Motor.eMotor;
};
\end{lstlisting}
\caption{Build Description}
\label{build}
\end{figure}

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/mrubyontecs.eps}
    \caption{System Model of existing mruby on TECS}
    \label{fig:mrubyontecs}
\end{figure}

\subsubsection{mruby-TECS Bridge}
There is a great difference between the execution time of mruby and C language.
According to  \cite{par:mrubyonTECS}, mruby programs are several hundreds times slower than C programs.
The execution of mruby bytecode on a RiteVM is not as efficient as that of C language.
Thus it is difficult to use mruby for all codes.

The use of Ruby on embedded devices provides the benefit of productivity and maintainability due to the ease to use and read.
On the other hands, it is necessary to implement parts of applications in C language in order to manipulate actuators and sensors, and also make a critical section of codes run quickly.

Figure \ref{fig:mruby_TECS_bridge} shows an example of use of an mruby-TECS bridge for controlling a motor.
The left side of BridgeMotor belongs to the mruby program.
The right side of BridgeMotor belongs to TECS component.
\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/mruby_TECS_bridge.eps}
    \caption{mruby-TECS Bridge}
    \label{fig:mruby_TECS_bridge}
\end{figure}

The mruby-TECS bridge generates two things.
One is a {\it celltype} to receive invocation from the mruby program.
The other is an mruby class that corresponds to the TECS component specified by the developers to invoke a C function from the mruby program.

@@

\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/system_model.eps}
    \caption{Detailed System Model of the proposed framework}
    \label{fig:system_model}
\end{figure}

\subsection{mruby Bytecode Loader Using Bluetooth}
\label{sec:mruby bytecode loader using Bluetooth}
\footnote{
mruby bytecode loader is intended to improve the development efficiency, therefore software developers should use it for development phase.
Complete software should be compiled and linked in the storage/ROM device beforehand.
}
This section describes an additional functionality of mruby on TECS, mruby bytecode loader using Bluetooth.
In the current system, the platform including mruby bytecodes are saved in a storage/ROM device.
Developers must rewrite the storage/ROM device every time the application programs are modified.
The OS on the target device should be also restarted.
It causes a low development efficiency to repeat that.
An mruby bytecode loader using Bluetooth makes the developers' burden decrease.
%The work that an SD card is in and out needs to be done just once in the beginning.
Developers should just once carry out the work to connect the storage/ROM device and start the OS.

mruby programs are consisted of mruby application and libraries.
mruby application is the main code which software developers should program.
mruby libraries are the codes defining the functions for application such as Ruby classes. 
The mruby bytecodes including mruby application and libraries can be sent and run.
However, it is also wasteful in terms of the size and time to send bytecodes because libraries are not frequently modified. 
The proposed framework provides the design that only mruby applications are sent and mruby libraries are preserved with the platform in the storage/ROM device beforehand.
Due to this design, RiteVMs can share the mruby libraries.
In addition, a RiteVM can use the own library that other RiteVMs should not use.

@@

% \begin{figure}[t]
%     \centering
%     \includegraphics[width=8cm,clip]{figure/bluetooth_loader.eps}
%     \caption{Development Flow in mruby bytecode loader using Bluetooth}
%     \label{fig:bluetooth_loader}
% \end{figure}

\subsubsection{Component of RiteVM with mruby bytecode loader using Bluetooth}
\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/component_bluetooth.eps}
    \caption{Component Diagram of mruby bytecode loader using Bluetooth}
    \label{fig:component_bluetooth}
\end{figure}
The proposed framework provides a RiteVM with mruby bytecode loader using Bluetooth as TECS component.
The component is an extension of the RiteVM component, which is described in \cite{par:mrubyonTECS}.
The component plays a role in receiving bytecodes via Bluetooth, and also manages a RiteVM configuration such as automatically generation the mruby library bytecode.
This generated bytecode is prepared beforehand in the storage/ROM device, and different from a bytecode transferred with Bluetooth.

Figure \ref{fig:component_bluetooth} shows a component diagram of MrubyTask1 and RiteVMBluetooth1 {\it cell}s.
The MrubyTask1 {\it cell} is a componentized task of the RTOS (TOPPERS/HRP2).
TOPPERS/HRP2 is described in \cite{url:HRP2}, \cite{par:hr-tecs}.
The RiteVMBluetooth1 is the component of RiteVM with mruby bytecode loader using Bluetooth.
A bytecode in the host is transferred and received at the top of the component.
In this framework, ZMODEM \cite{par:zmodem} is used as a binary transfer protocol.

\begin{figure}[t]
    \centering
    \includegraphics[width=8cm,clip]{figure/control_flow.eps}
    \caption{Process Flow of mruby bytecode loader using Bluetooth}
    \label{fig:control_flow}
\end{figure}

\begin{figure}[t]
\centering
\begin{lstlisting}
/* tRiteVMBluetooth.cdl */
celltype tRiteVMBluetooth{
    entry sTaskBody eMrubyBody;
    [optional] call sEventflag cEventflag[];
    [optional] call sSemaphore cSemaphore;
    attr{
      [omit]char_t *mrubyLib;
      char_t *irepLib = 
                    C_EXP("&$cell_global$_irep");
      uint32_t irepAppSize = 
                    C_EXP( BUFFER_SIZE );
      FLGPTN setptn;
    };
    var{
        mrb_state *mrb;
        mrbc_context *context;
        [size_is(irepAppSize)] uint8_t *irepApp;
    };
};
\end{lstlisting}
\caption{Celltype Description for RiteVM with mruby bytecode loader using Bluetooth}
\label{celltype_mrubybluetooth}
\end{figure}

Figure \ref{fig:control_flow} shows the process flow of executing mruby program in the component of RiteVM with mruby bytecode loader using Bluetooth such as RiteVMBluetooth1.
The concrete main code of tRiteVMBluetooth is shown in Figure \ref{maincode_mrubybluetooth}.

First, the mruby bytecode loader receives the mruby application bytecode transferred from the host (Figure \ref{fig:control_flow}:A, Figure \ref{maincode_mrubybluetooth}:6-7).
The bytecode is stored in the component variable such as VAR\_irepApp shown in Figure \ref{celltype_mrubybluetooth}.
This process is exclusively carried out by the semaphore.

Second, pointers of {\it mrb\_state} and {\it mrbc\_context}, and mruby-TECS bridges are initialized (Figure \ref{fig:control_flow}:B, Figure \ref{maincode_mrubybluetooth}:9-14).
{\it mrb\_state} is a set of states and global variables used in mruby.
Synchronization of multiple tasks is performed in this processing phase.
The RiteVM that ended up here waits for the other RiteVM to finish loading and initialization.

Third, the RiteVM reads the bytecode of mruby libraries (Figure \ref{fig:control_flow}:C, Figure \ref{maincode_mrubybluetooth}:17-19).
mruby libraries are a set of Ruby classes such as motor class and sensor class.
For example, motor class defines methods to rotate and stop a motor.
The tRiteVMBluetooth {\it cell} has attributes as shown in Figure \ref{celltype_mrubybluetooth}.
The {\it mrubyLib} indicates the program files of mruby libraries.
{\it [omit]} is only used for the TECS generator, thus the attribute, {\it mrubyLib}, does not consume memory.
The {\it irepLib} is the pointer of the array stored the bytecode of mruby libraries.
In short, the bytecode of mruby libraries is stored as an attribute of the component when compiling for the first time.

Fourth, the RiteVM reads the bytecode of the mruby application transferred with Bluetooth (Figure \ref{fig:control_flow}:D, Figure \ref{maincode_mrubybluetooth}:20-22).
The mruby application bytecode is stored in an array of {\it irepApp}.
The array is different from that of holding the mruby library bytecode.
Two bytecodes are read separately in the RiteVM.

Finally, the mruby task runs (Figure \ref{fig:control_flow}:E, Figure \ref{maincode_mrubybluetooth}:20-22).
When the mruby application is modified, only the bytecode of the modified application should be transferred.
mruby libraries need not be touched because libraries are not normally changed.
